---
title: "Food Security Indicator Plausbility Checks"
author: "William Mcfall - WFP South Sudan RAM Officer"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: espresso
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: yes
      smooth_scroll: no
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
---


```{r script_setup, include=FALSE, fig.align = "center", fig.width= 16, fig.height= 11, fig.asp= 0.618, out.width="100%"}

############
### Removes Objects in Memory
############

rm(list = ls())

############
### Establishes Default Settings for Knitr
############
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	cache.path = "cache/html/" # set to ensure that cache directory functions properly... change path if output type changes
)

knitr::opts_knit$set(root.dir = 'XXX') 

############
### Establishes the FSNMS Project Working Directory
############

setwd('XXX')

############
### Loads the Required Libraries
############

library(gtsummary)
library(readxl)
library(openxlsx)
library(dplyr)
library(tidyr)
library(mosaic)
library(tidyverse)
library(modelsummary)
library(kableExtra) 
library(foreign)
library(arsenal)
library(desctable)
library(survey)
library(srvyr) 
library(naniar)
library(rmarkdown)
library(haven)
library(labelled)
library(sjlabelled)
# library(xlsx)
library(rlang)
library(digitTests)
library(skimr)
library(ggrepel)
library(janitor)
library(skimr)
library(skimr)
library(ggmap)
library(RM.weights)
library("gt")
library("stringr")
library(sf)
library(rgdal)
library(FRK)


```

```{r data_import}

########################################################################
### Establishes the FSNMS Local File Path
########################################################################

file_path <- "./Data Input/cleaned_raw_data.xlsx"

########################################################################
### Creates Household-Level Food Security Dataset
########################################################################

## imports household-level food security dataset
fsnms_data <- read_xlsx(file_path, sheet="data")

########################################################################
### Creates Error Flag Variables Based on Plausibility Assumptions
########################################################################

# calculates standard deviation and mean food expenditure
sd_food_exp <- sd(fsnms_data$HHExpFood_1M)
mean_food_exp <- mean(fsnms_data$HHExpFood_1M)

## creates error flag variables indicating if the data collection was fast, slow, or negative; rCSI was high; FCS was zero, low, or high; food expenditure was zero, low, or high
fsnms_data %>% mutate(errorflag_duration_fast = ifelse(duration_minutes <= 30 & duration_minutes >= 0, 1,0),
                      errorflag_duration_slow = ifelse(duration_minutes > 180, 1, 0),
                      errorflag_duration_negative = ifelse(duration_minutes < 0, 1, 0),
                      errorflag_rcsi_high = ifelse(rCSI >= 45, 1, 0),
                      errorflag_fcs_zero = ifelse(FCS == 0, 1, 0),
                      errorflag_fcs_low = ifelse(FCS < 10, 1, 0),
                      errorflag_fcs_high = ifelse(FCS > 100, 1, 0),
                      errorflag_fcs_staple_zero = ifelse(FCSStap == 0, 1, 0),
                      errorflag_food_exp_high = ifelse(HHExpFood_1M >= (mean_food_exp + 2*sd_food_exp), 1,0),
                      errorflag_food_exp_low = ifelse(HHExpFood_1M <= (mean_food_exp - 2*sd_food_exp), 1,0),
                      errorflag_food_exp_zero = ifelse(HHExpFood_1M == 0, 1,0)) -> fsnms_data
